Brief Announcement: Tight Bounds for Online Balanced Repartitioning

### 2020-03-27 paper/paper.tex
Sketch of 2k-1 lb

We partition nodes into $3$ sets: $A$, $B$ and $C$.
Sets has the following cardinality: $|A| = k-1$, $|B| = k-1$, and $|C| = 2$.
We arbitratry assign all elements of the universe $U$ to nodes from $A\cup B\cup C$.
Note that in every configuration, at least one of these sets is split, i.e., is not completely contained in one cluster.

>>>>

\begin{theorem}
  Every deterministic algorithm for BRP with $\ell=2$ is at least $c$-competitive, where $c=2k-1$.
\end{theorem}

\begin{proof}
We partition nodes into $3$ sets: $A$, $B$ and $C$.
Sets has the following cardinality: $|A| = k-1$, $|B| = k-1$, and $|C| = 2$.
We arbitrarily assign all elements of the universe $U$ to nodes from $A\cup B\cup C$.

Note that in every configuration, at least one of these sets is split, i.e., is not completely contained in one cluster.
Intuitively, if a node is split from other nodes of any of sets $A,B,C$, then it is outside of the cache.

Similarly to the lower bound TODO ref $\Omega(kl)$, we construct components of nodes, and issue the requests if the algorithm splits the component.
(In the current proof, the components produced are oblivious to the configuration of the algorithm, yet the requests issued depend on if the algorithm splits the components. In $\Omega(kl)$, the components were constructed upon the actions of the algorithm, too.)

Now, we show how to simulate an algorithm for the paging problem using the constructed BRP instance.
Fix any algorithm for the paging problem.
If a request to element $e$ comes, we identify its corresponding node $v_e$, and the set $S$ from $A,B,C$ that contains it.
Then, the adversary produces a component $S$, and removes it right after the algorithm collocates the whole $S$ on one cluster.

Now we bound the costs of the algorithm for any sequence of paging requests transformed the way described above.
For any request, the algorithm pays $0$ if it has $S$ collocated, and otherwise it pays at least $2\alpha$ for at least one migration needed to achieve state where $S$ is collocated.

Now we construct an offline algorithm OPT.
OPT only resides in three types of configurations:
\begin{enumerate}
  \item With set $A$ collocated in one cluster, set $B$ collocated in the other, and set $B$ having one node in one cluster and the other node in the second
  \item With set $A$ collocated in one cluster, with a single node from set $B$; on the second cluster, we have both nodes from set $C$, and $k-2$ nodes from set $B$
  \item the symmetric configuration to the previous one, with set $A$ split, and sets $B$ and $C$ collocated
\end{enumerate}
To transition between any of these configuration types, a single exchange is needed, hence OPT pays $2\alpha$ for it.
Within each type of configuration, OPT may exchange the split node with any node from his set for $2\alpha$.

OPT observes the actions of the algorithm LFD (Least Forward Distance) for the caching problem, and performs adequate reconfigurations within feasible configurations described in the previous paragraph.
Precisely, if LFD chooses to evict the element $e$ from the cache, OPT chooses the configuration with the corresponding node $v_e$ split, and pays $2\alpha$ for it.

The well-known property of LFD is that it pays $1$ only after $\mathbb{C}$ different pages are revealed to the algorithm, where $\mathbb{C}$ is the size of its cache.
In our case, $\mathbb{C} = 2k-1$, and hence the algorithm pays at least $2\alpha$ for each (transformed to BRP) paging request, and OPT pays $2\alpha$ every $2k-1$ paging requests.
Hence, the claim holds.
\end{proof}
